Supporting Information

### 2020-07-13 manuscript/manuscript.Rmd
Updated map reference sample, some changes to SI, moving map reference sample selection out of analyses to own vignette

# Discussion
- Left unquantified is how much training on base maps matters
- Image quality:
    - why mean-weighting not median composite (artifacts) (maybe methods)
    - impact on accuracy
- Assessment of segmentation quality
- Next steps: 
    - Annually updated cropland maps
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.

>>>>

# Discussion
- Left unquantified is how much training on base maps matters
- Image quality:
    - why mean-weighting not median composite (artifacts) (maybe methods)
    - impact on accuracy
- Assessment of segmentation quality
- Next steps: 
    - Annually updated cropland maps
    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.

### 2020-07-13 manuscript/si.Rmd
Updated map reference sample, some changes to SI, moving map reference sample selection out of analyses to own vignette

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

@@

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library(knitr)
library(citr)

@@

### Accuracy assessment

@@

>>>>

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: false
spacing: doublespacing
footerdate: yes
abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

@@

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(knitr)
library(citr)
library(tidyverse)
library(sf)

@@

@@

### Accuracy assessment
We designed and implemented a map accuracy assessment protocol following procedures summarized by @StehmanKeyissuesrigorous2019. This entailed the creation of a map reference sample, which first entailed designing a sample, and then designing how the sample response would be collected.

#### Map reference sample design
A reference sample should ideally have some randomization involved in the sample selection protocol, and the probabilities for a sample being included should be provided.  There are multiple sample designs that can be chosen, but we employed a stratified sampling design, using the segmented field boundaries to define the strata for field/non-field. To create the sample, we first extracted the centroids for each of the sample grid cells in Ghana. We then intersected the centroid points with the field segments, assigned a class of 1 (cropland) where points intersected a field, and 0 where they didn't (non-cropland). We then removed from this set of points all those that corresponded to model training, validation, or training reference sites. We then extracted a random sample from both the cropland and non-cropland points. To determine the sample size of each, we specified a desired confidence interval using the following formula [@StehmanKeyissuesrigorous2019]:

$$n = \frac{z^2p(1-p)}{d^2}$$

Where p is the estimated probability (or mapped class accuracy), and d is the size of the margin of error (1/2 the confidence interval). We selected a d value of 0.03 and assumed that the user's accuracy of the field class would be 0.75 and that of the non-cropland class would be 0.8, returning sample sizes of 800 and 683, respectively. The distribution of the resulting map reference sample is shown in Figure 1.

```{r, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
africa <- st_read(
  system.file("extdata/africa.geojson", package = "activemapper"), quiet = TRUE,
)
aois <- st_read(system.file("extdata/aois.geojson", package = "activemapper"), 
                quiet = TRUE)
africa_cont <- africa %>% st_union %>% st_sf
ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)

ref_sample <- readr::read_csv(
  here::here("external/reference/map_reference_sample.csv")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

p <- ggplot() + 
  # geom_sf(data = ghana) +
  geom_sf(data = aois) + 
  geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) +
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Non-cropland", "Cropland"), name = "") + 
  theme_light()
p
```

#### Response design
##### Stage 1
We collected the map reference sample on a separate instance of `labeller` set up for the purpose (`validator.crowdmapper.org`). For the sampling unit, we selected a rectangular polygon of ~0.1 ha (0.0002866$^\circ$ resolution), centered on the centroid of each grid cell selected for map reference sample.  Four classes were established for the validation: **cropland**, **non-cropland**, **uncertain but likely cropland**, and **uncertain but likely non-cropland**. The latter two classes were designed to capture information related to swidden dynamics, following the rationale that uncertainty and time since last cropping are likely to be positively correlated. This uncertainty also captures information about the inherent difficulty of the mapping task. Samples were collected by visually interpreting the overlays of PlanetScope composites presented in `labeller`, following the same interpretation protocols used by the labelling team, with the exception that the polygons placed were square and of 0.0002866$^\circ$ resolution.

The map reference sample was first placed and defined in an initial stage. In this stage, the class of the initial square polygon was determined, and then the polygon was adjusted if it overlapped more than one class. In sites where the initial map reference grid fell on one of the two uncertain classes, a second polygon was placed in the nearest location where a definite cropland or non-cropland interpretation could be made, with the choice determined by whether the uncertain class was more likely cropland or not cropland. The uncertain class was only used for providing insight into fallow cycles, while the second, more certain polygon was used in a second stage of the map reference sample, which was given to multiple independent interpreters (see below).

The following set of rules were followed in this first, sample placing phase (undertaken by Lyndon Estes):

1. When determining the class corresponding with the initial location of the target grid, IF:
  - More than half of the target falls within what appears to be a clear arable crop field, then assign it to **cropland**. 
  - More than half falls in what is clearly not a field, then assign it to **non-cropland**. 
  - More than half falls in a location where it is harder to tell whether it is cropland or non-cropland, determine whether it is more likely a crop field or not a crop field, and then assign either **uncertain but likely cropland**, or **uncertain but likely non-cropland**.

2. After determining the class, IF:
  a. The target polygon is contained entirely within a single clear class AND
      - The class is either **cropland** or **non-cropland** THEN:
        - Digitize a point within the center of the target box, assign appropriate class, and complete the assignment
      - The class belongs to one of the two **uncertain** classes THEN:

        - Digitize a square aligned exactly with the initial target polygon, and choose the class **cropland** or **non-cropland** (following the most likely interpretation) and move the new polygon to the nearest location where it can be contained entirely within that clear **cropland** or **non-cropland** class;

        - Digitize a point in the middle of the target polygon, and assign it the appropriate label: **cropland** or **non-cropland**.

  b. The target polygon partially overlaps a second class AND:
      - That class is either **cropland** or **non-cropland** THEN:

        - Digitize a square aligned exactly with the target polygon, assign the appropriate class, and move the box to the nearest location (shortest possible distance) where it can be contained entirely within that same class.

      - That class is **uncertain** THEN:

        - Digitize a square aligned exactly with the target polygon, assign the appropriate **uncertain** class, then move the box to the nearest location (shortest possible distance) where it can be contained entirely within that same **uncertain** class.
        - Digitize a second square aligned exactly with the initial target polygon, and choose the class **cropland** or **non-cropland** (following the most likely interpretation) and move the new polygon to the nearest location where it can be contained entirely within that clear **cropland** or **non-cropland** class.

We then 
  - Definition of spatial assessment
  - Definition of classes
  - Reference source
  - Specific information collected from each source
  - Rules for reference class labels
  - How agreement between map and reference classes is defined

##### Stage 2
